---
title: "emtdata"
author: "Malvika D. Kharbanda"
date: "`r BiocStyle::doc_date()`"
output:
  prettydoc::html_pretty:
    theme: hpstr
    toc: yes
    toc_depth: 2
    number_sections: yes
    fig_caption: yes
    df_print: paged
vignette: >
  %\VignetteIndexEntry{emtdata}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# emtdata

This package can be installed using the code below:

```{r install}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("emtdata")
```

# Download data from the emtdata R package

Data in this package can be downloaded using the `ExperimentHub` interface as shown below. To download the data, we first need to get a list of the data available in the `emtdata` package and determine the unique identifiers for each data. The `query()` function assists in getting this list.

```{r load-packages, message=FALSE}
library(emtdata)
library(ExperimentHub)
```

```{r get-emtdata}
eh = ExperimentHub()
query(eh , 'emtdata')
```

Data can then be downloaded using the unique identifier.

```{r download-emtdata-cursons2018-id}
eh[['EH5440']]
```

Alternatively, data can be downloaded using object name accessors in the `emtdata` package as below:

```{r download-emtdata-cursons2018-accessor}
#metadata are displayed
cursons2018_se(metadata = TRUE)
#data are loaded
cursons2018_se()
```

# Accessing SummarizedExperiment object

```{r access-se}
cursons2018_se = eh[['EH5440']]

#read counts
assay(cursons2018_se)

#genes
rowData(cursons2018_se)

#sample information
colData(cursons2018_se)
```

# Exploratory analysis and visualization 

```{r}
library("DESeq2")
cursons2018_dds <- DESeqDataSet(cursons2018_se, design = ~ Subline + Treatment)
cursons2018_rld <- rlog(cursons2018_dds)
cursons2018_sampleDists <- dist( t( assay(cursons2018_rld) ) )
```

## PCA
```{r}
cursons2018_data <- plotPCA(cursons2018_rld, intgroup = c( "Treatment", "Subline"), returnData=TRUE)
cursons2018_percentVar <- round(100 * attr(cursons2018_data, "percentVar"))
library("ggplot2")
ggplot(data, aes(PC1, PC2, color=Treatment, shape=Subline)) + geom_point(size=3) +
  xlab(paste0("PC1: ",cursons2018_percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",cursons2018_percentVar[2],"% variance"))
```

## MDS
```{r}
cursons2018_sampleDistMatrix <- as.matrix( cursons2018_sampleDists )
cursons2018_mds <- data.frame(cmdscale(cursons2018_sampleDistMatrix))
cursons2018_mds <- cbind(cursons2018_mds, as.data.frame(colData(cursons2018_rld)))
ggplot(cursons2018_mds, aes(X1,X2,color=Treatment,shape=Subline)) + geom_point(size=3)
```

# Session information

```{r sessionInfo}
sessionInfo()
```
